---
title: "Unit 10 activities"
author: "Guilherme Amorim"
date: "2024-10-16"
output: html_document
---

# Unit 10 data activities

1. Find out correlation between systolic and diastolic BP.
2. Produce a scatter plot between systolic and diastolic BP.

## R

### Initial setup

```{r, include=T, message=F}
# initial setup
library(haven)
library(skimr)
library(tidyverse)
library(here)
```

### Data load
```{r}
Health_Data<-read_sav(here("Datasets/Health Data.sav"))

```

### Correlation

```{r}
cor.test(Health_Data$sbp, Health_Data$dbp)

```

r (Pearson's correlation coefficient)= 0.846, p-value <0.001, indicating a statistically significant strong correlation between systolic and diastolic blood pressure

### Scatterplot

```{r}

Health_Data%>%
  ggplot(aes(sbp, dbp))+
  geom_point()


```

## Python

### Initial setup


```{python}
# first install packages (from terminal)

# pip3 install numpy

# pip3 install pandas

# pip3 install matplotlib 

# pip3 install pyreadstat 

import pandas as pd

import numpy as np 

import matplotlib.pyplot as plt

from scipy import stats
```



### Data load

```{python}

Health_Data_python = pd.read_spss("C:/Users/guilhermep/Documents/PgDip/Coding/Module 2/pgdip_module2_practice/Datasets/Health Data.sav")


```

### Correlation

```{python}

stats.pearsonr(Health_Data_python.sbp, Health_Data_python.dbp)

```

### Scatterplot

```{python}
plt.scatter(Health_Data_python.sbp, Health_Data_python.dbp, alpha=0.5)
plt.show()
```